Tropical Andes Functional diversity for plants and Frugivores

### 2023-08-05 code/L2/L2_2_TropicalAndes_funcdiv.Rmd
add plotting code fix mismatched name issue

```{r}
# getting sp_faxes_coord_plants and subset_matrix_plants names to match
sp_faxes_coord_plants_sub <- sp_faxes_coord_plants[ , c("PC1", "PC2", "PC3", "PC4")]
sp_faxes_coord_plants_sub <- as.data.frame(sp_faxes_coord_plants_sub)
sp_faxes_coord_plants_sub_names <- row.names(sp_faxes_coord_plants_sub)
subset_matrix_plant <- as.data.frame(subset_matrix_plant)
subset_matrix_plant_names <- colnames(subset_matrix_plant)
subset_matrix_plant_sub <- subset_matrix_plant %>% filter(names(subset_matrix_plant) %in% sp_faxes_coord_plants_sub_names)
```

```{r}
#computing FD
# The number of species per assemblage has to be higher or equal to the number of traits
alpha_fd_indices_plant <- mFD::alpha.fd.multidim(
  sp_faxes_coord   = sp_faxes_coord_plants[ , c("PC1", "PC2", "PC3", "PC4")],
  asb_sp_w         = subset_matrix_plant,
  ind_vect         = c("fide", "fdis", "fmpd", "fnnd", "feve", "fori", "fspe"), #did not run fdiv or fric (slows it down)
  scaling          = TRUE,
  check_input      = TRUE,
  details_returned = TRUE)

alpha_fd_indices_frugivore <- mFD::alpha.fd.multidim(
  sp_faxes_coord   = sp_faxes_coord_frugivore[ , c("PC1", "PC2", "PC3", "PC4")],
  asb_sp_w         = subset_matrix_frugivore,
  ind_vect         = c("fide", "fdis", "fmpd", "fnnd", "feve", "fori", "fspe"), #did not run fdiv or fric (slows it down)
  scaling          = TRUE,
  check_input      = TRUE,
  details_returned = TRUE)

@@

#plot
plots_alpha_plant <- mFD::alpha.multidim.plot(
  output_alpha_fd_multidim = alpha_fd_indices_plant,
  plot_asb_nm              = c("cell40", "cell200"),
  ind_nm                   = c("fdis", "fide", "fnnd"
                        , "fori", "fspe"),
  faxes                    = NULL,
  faxes_nm                 = NULL,
  range_faxes              = c(NA, NA),
  color_bg                 = "grey95",
  shape_sp                 = c(pool = 3, asb1 = 21, asb2 = 21),
  size_sp                  = c(pool = 0.7, asb1 = 1, asb2 = 1),
  color_sp                 = c(pool = "grey50", asb1 = "#1F968BFF", asb2 = "#DCE319FF"),
  color_vert               = c(pool = "grey50", asb1 = "#1F968BFF", asb2 = "#DCE319FF"),
  fill_sp                  = c(pool = NA, asb1 = "#1F968BFF", asb2 = "#DCE319FF"),
  fill_vert                = c(pool = NA, asb1 = "#1F968BFF", asb2 = "#DCE319FF"),
  color_ch          
  = c(pool = NA, asb1 = "#1F968BFF", asb2 = "#DCE319FF"),
  fill_ch                  = c(pool = "white", asb1 = "#1F968BFF", asb2 = "#DCE319FF"),
  alpha_ch                 = c(pool = 1, asb1 = 0.3, asb2 = 0.3),
  shape_centroid_fdis      = c(asb1 = 22,  asb2 = 24),
  shape_centroid_fdiv      = c(asb1 = 22,  asb2 = 24),
  shape_centroid_fspe      = 23,
  color_centroid_fspe      = "black",
  size_sp_nm               = 3, 
  color_sp_nm              = "black",
  plot_sp_nm               = NULL,
  fontface_sp_nm           = "plain",
  save_file                = FALSE,
  check_input              = TRUE) 
plots_alpha_frugivore <- mFD::alpha.multidim.plot(
  output_alpha_fd_multidim = alpha_fd_indices_frugivore,
  plot_asb_nm              = c("cell40", "cell200"),
  ind_nm                   = c("fdis", "fide", "fnnd"
                        , "fori", "fspe"),
  faxes                    = NULL,
  faxes_nm                 = NULL,
  range_faxes              = c(NA, NA),
  color_bg                 = "grey95",
  shape_sp                 = c(pool = 3, asb1 = 21, asb2 = 21),
  size_sp                  = c(pool = 0.7, asb1 = 1, asb2 = 1),
  color_sp                 = c(pool = "grey50", asb1 = "#1F968BFF", asb2 = "#DCE319FF"),
  color_vert               = c(pool = "grey50", asb1 = "#1F968BFF", asb2 = "#DCE319FF"),
  fill_sp                  = c(pool = NA, asb1 = "#1F968BFF", asb2 = "#DCE319FF"),
  fill_vert                = c(pool = NA, asb1 = "#1F968BFF", asb2 = "#DCE319FF"),
  color_ch          
  = c(pool = NA, asb1 = "#1F968BFF", asb2 = "#DCE319FF"),
  fill_ch                  = c(pool = "white", asb1 = "#1F968BFF", asb2 = "#DCE319FF"),
  alpha_ch                 = c(pool = 1, asb1 = 0.3, asb2 = 0.3),
  shape_centroid_fdis      = c(asb1 = 22,  asb2 = 24),
  shape_centroid_fdiv      = c(asb1 = 22,  asb2 = 24),
  shape_centroid_fspe      = 23,
  color_centroid_fspe      = "black",
  size_sp_nm               = 3, 
  color_sp_nm              = "black",
  plot_sp_nm               = NULL,
  fontface_sp_nm           = "plain",
  save_file                = FALSE,
  check_input              = TRUE)

@@

```{r}
## Turning the FD values into a raster

@@

plant_fd_sp <- data.frame(subset_coords_plant_sp, fdis_plant)
plant_fd_sp <- data.frame(subset_coords_plant_sp, fdis_plant)

# Convert the dataframe to sf format
spatial_fdis_bird <- st_as_sf(bird_fd_sp, coords = c("Longitude.x.", "Latitude.y."))

# Rasterize the sf data to create the FD raster
fd_raster_bird <- rasterize(spatial_fdis_bird, empty_raster)
writeRaster(fd_raster_bird$fdis_bird, "FD_birds_TA.tif", format="GTiff",overwrite=T)
writeRaster(fd_raster_bird$ID, "FD_birds_TA_ID.tif", format="GTiff")

@@

>>>>

```{r}
# getting sp_faxes_coord_plants and subset_matrix_plants names to match
sp_faxes_coord_plants_sub <- sp_faxes_coord_plants[ , c("PC1", "PC2", "PC3", "PC4")]
## check number of species names
nrow(sp_faxes_coord_plants_sub) # 54
ncol(subset_matrix_plant) #82

sp_faxes_coord_plants_sub_names <- row.names(sp_faxes_coord_plants_sub)
subset_matrix_plant_names <- colnames(subset_matrix_plant)

sp_faxes_coord_plants_sub_df <- as.data.frame(sp_faxes_coord_plants_sub)
subset_matrix_plant_df <- as.data.frame(subset_matrix_plant)

subset_matrix_plant <- subset_matrix_plant_df[, which((names(subset_matrix_plant_df) %in% sp_faxes_coord_plants_sub_names)==TRUE)]

ncol(subset_matrix_plant)
subset_matrix_plant <- as.matrix(subset_matrix_plant)
```

```{r}
# match frugivore names
sp_faxes_coord_frugivore_sub <- sp_faxes_coord_frugivore[ , c("PC1", "PC2", "PC3", "PC4")]

## check number of species names
nrow(sp_faxes_coord_frugivore_sub) #718
ncol(subset_matrix_frugivore) #712

sp_faxes_coord_frugivore_sub_names <- row.names(sp_faxes_coord_frugivore_sub)
subset_matrix_frugivore_names <- colnames(subset_matrix_frugivore)

sp_faxes_coord_frugivore_sub_names <- chartr(".", " ", sp_faxes_coord_frugivore_sub_names)
sp_faxes_coord_frugivore_sub <- as.data.frame(sp_faxes_coord_frugivore_sub)
row.names(sp_faxes_coord_frugivore_sub) <- sp_faxes_coord_frugivore_sub_names

sp_faxes_coord_frugivore_sub <- sp_faxes_coord_frugivore_sub[ which((row.names(sp_faxes_coord_frugivore_sub) %in% subset_matrix_frugivore_names)==TRUE), ]

nrow(sp_faxes_coord_frugivore_sub)
sp_faxes_coord_frugivore_sub <- as.matrix(sp_faxes_coord_frugivore_sub)
```

```{r}
#computing FD
# The number of species per assemblage has to be higher or equal to the number of traits
alpha_fd_indices_plant <- mFD::alpha.fd.multidim(
  sp_faxes_coord   = sp_faxes_coord_plants_sub,
  asb_sp_w         = subset_matrix_plant,
  ind_vect         = c("fide", "fdis", "fmpd", "fnnd", "feve", "fori", "fspe"), #did not run fdiv or fric (slows it down)
  scaling          = TRUE,
  check_input      = TRUE,
  details_returned = TRUE)

alpha_fd_indices_frugivore <- mFD::alpha.fd.multidim(
  sp_faxes_coord   = sp_faxes_coord_frugivore_sub,
  asb_sp_w         = subset_matrix_frugivore,
  ind_vect         = c("fide", "fdis", "fmpd", "fnnd", "feve", "fori", "fspe"), #did not run fdiv or fric (slows it down)
  scaling          = TRUE,
  check_input      = TRUE,
  details_returned = TRUE)

@@

#plot
plots_alpha_plant <- mFD::alpha.multidim.plot(
  output_alpha_fd_multidim = alpha_fd_indices_plant,
  plot_asb_nm              = c("cell4", "cell20"),
  ind_nm                   = c("fdis", "fide", "fnnd"
                        , "fori", "fspe"),
  faxes                    = NULL,
  faxes_nm                 = NULL,
  range_faxes              = c(NA, NA),
  color_bg                 = "grey95",
  shape_sp                 = c(pool = 3, asb1 = 21, asb2 = 21),
  size_sp                  = c(pool = 0.7, asb1 = 1, asb2 = 1),
  color_sp                 = c(pool = "grey50", asb1 = "#1F968BFF", asb2 = "#DCE319FF"),
  color_vert               = c(pool = "grey50", asb1 = "#1F968BFF", asb2 = "#DCE319FF"),
  fill_sp                  = c(pool = NA, asb1 = "#1F968BFF", asb2 = "#DCE319FF"),
  fill_vert                = c(pool = NA, asb1 = "#1F968BFF", asb2 = "#DCE319FF"),
  color_ch          
  = c(pool = NA, asb1 = "#1F968BFF", asb2 = "#DCE319FF"),
  fill_ch                  = c(pool = "white", asb1 = "#1F968BFF", asb2 = "#DCE319FF"),
  alpha_ch                 = c(pool = 1, asb1 = 0.3, asb2 = 0.3),
  shape_centroid_fdis      = c(asb1 = 22,  asb2 = 24),
  shape_centroid_fdiv      = c(asb1 = 22,  asb2 = 24),
  shape_centroid_fspe      = 23,
  color_centroid_fspe      = "black",
  size_sp_nm               = 3, 
  color_sp_nm              = "black",
  plot_sp_nm               = NULL,
  fontface_sp_nm           = "plain",
  save_file                = FALSE,
  check_input              = TRUE)

plots_alpha_frugivore <- mFD::alpha.multidim.plot(
  output_alpha_fd_multidim = alpha_fd_indices_frugivore,
  plot_asb_nm              = c("cell4", "cell20"),
  ind_nm                   = c("fdis", "fide", "fnnd"
                        , "fori", "fspe"),
  faxes                    = NULL,
  faxes_nm                 = NULL,
  range_faxes              = c(NA, NA),
  color_bg                 = "grey95",
  shape_sp                 = c(pool = 3, asb1 = 21, asb2 = 21),
  size_sp                  = c(pool = 0.7, asb1 = 1, asb2 = 1),
  color_sp                 = c(pool = "grey50", asb1 = "#1F968BFF", asb2 = "#DCE319FF"),
  color_vert               = c(pool = "grey50", asb1 = "#1F968BFF", asb2 = "#DCE319FF"),
  fill_sp                  = c(pool = NA, asb1 = "#1F968BFF", asb2 = "#DCE319FF"),
  fill_vert                = c(pool = NA, asb1 = "#1F968BFF", asb2 = "#DCE319FF"),
  color_ch          
  = c(pool = NA, asb1 = "#1F968BFF", asb2 = "#DCE319FF"),
  fill_ch                  = c(pool = "white", asb1 = "#1F968BFF", asb2 = "#DCE319FF"),
  alpha_ch                 = c(pool = 1, asb1 = 0.3, asb2 = 0.3),
  shape_centroid_fdis      = c(asb1 = 22,  asb2 = 24),
  shape_centroid_fdiv      = c(asb1 = 22,  asb2 = 24),
  shape_centroid_fspe      = 23,
  color_centroid_fspe      = "black",
  size_sp_nm               = 3, 
  color_sp_nm              = "black",
  plot_sp_nm               = NULL,
  fontface_sp_nm           = "plain",
  save_file                = FALSE,
  check_input              = TRUE)

@@

# Turning the FD values into a raster
```{r}

@@

plant_fd_sp <- data.frame(subset_coords_plant_sp, fdis_plant)
frugivore_fd_sp <- data.frame(subset_coords_frugivore_sp, fdis_frugivore)

# Convert the dataframe to sf format
spatial_fdis_plant <- st_as_sf(plant_fd_sp, coords = c("Longitude.x.", "Latitude.y."))
spatial_fdis_frugivore <- st_as_sf(frugivore_fd_sp, coords = c("Longitude.x.", "Latitude.y."))

@@

```{r}
# Rasterize the sf data to create the FD raster
fd_raster_plant <- rasterize(spatial_fdis_plant, empty_raster)
writeRaster(fd_raster_plant$fdis_plant, filename = file.path(output_path,"FD_plant_TA.tif"), format="GTiff",overwrite=T)
writeRaster(fd_raster_plant$ID, filename = file.path(output_path,"FD_plant_TA_ID.tif"), format="GTiff")

fd_raster_frugivore <- rasterize(spatial_fdis_frugivore, empty_raster)
writeRaster(fd_raster_frugivore$fdis_frugivore, filename = file.path(output_path,"FD_frugivore_TA.tif"), format="GTiff",overwrite=T)
writeRaster(fd_raster_frugivore$ID, filename = file.path(output_path,"FD_frugivore_TA_ID.tif"), format="GTiff")
```

# Functional diversity plots
```{r}
# set crs of sf objects
spatial_fdis_plant <- spatial_fdis_plant %>% st_set_crs(4326)
spatial_fdis_frugivore <- spatial_fdis_frugivore %>% st_set_crs(4326)
```

```{r}
# Raster to raster point
spec_FD_raster_plant_point <- rasterToPoints(fd_raster_plant, spatial = TRUE)
spec_FD_raster_frugivore_point <- rasterToPoints(fd_raster_frugivore, spatial = TRUE)
# Convert to a 'conventional' dataframe
fd_raster_plant_point <- data.frame(spec_FD_raster_plant_point)
fd_raster_frugivore_point <- data.frame(spec_FD_raster_frugivore_point)
```

```{r}
plantgridFDTA <-
  ggplot() +
  geom_sf(data = worldMap, fill = "white")+
  geom_sf(data = TApoly, fill = "grey", size = 0.1) +
  geom_tile(data= fd_raster_plant_point, aes(x=x, y=y, fill=fdis_plant)) +
  scale_fill_scico(palette = "romaO", direction = 1) +
  coord_sf(xlim = c(-85, -54), ylim = c(-24, 14), expand = FALSE) +
  theme(
    plot.background = element_rect(fill = "#f1f2f3"),
    panel.background = element_rect(fill = "lightblue"),
    panel.grid = element_blank(),
    line = element_blank(),
    rect = element_blank()
  ) +
  annotation_scale(location = "bl", width_hint = 0.5) +
  annotation_north_arrow(location = "bl", which_north = "true", 
  pad_x = unit(0.3, "in"), pad_y = unit(0.3, "in"), style = north_arrow_fancy_orienteering) +
  labs(title = "Tropical Andes Plant Functional Dispersion", x = "Longitude", y = "Latitude", fill = "FD")
plantgridFDTA
```

```{r}
frugivoregridFDTA <-
  ggplot() +
  geom_sf(data = worldMap, fill = "white")+
  geom_sf(data = TApoly, fill = "grey", size = 0.1) +
  geom_tile(data= fd_raster_frugivore_point, aes(x=x, y=y, fill=fdis_frugivore)) +
  scale_fill_scico(palette = "romaO", direction = 1) +
  coord_sf(xlim = c(-85, -54), ylim = c(-24, 14), expand = FALSE) +
  theme(
    plot.background = element_rect(fill = "#f1f2f3"),
    panel.background = element_rect(fill = "lightblue"),
    panel.grid = element_blank(),
    line = element_blank(),
    rect = element_blank()
  ) +
  annotation_scale(location = "bl", width_hint = 0.5) +
  annotation_north_arrow(location = "bl", which_north = "true", 
  pad_x = unit(0.3, "in"), pad_y = unit(0.3, "in"), style = north_arrow_fancy_orienteering) +
  labs(title = "Tropical Andes Frugivore Functional Dispersion", x = "Longitude", y = "Latitude", fill = "FD")
frugivoregridFDTA
```

# Extract cell values
```{r}
plantgridFDTA <- plantgridFDTA[["data"]]
frugivoregridFDTA <- frugivoregridFDTA[["data"]]
```

```{r}
plot(spatial_fdis_plant$fdis_plant, spatial_fdis_frugivore$fdis_frugivore,
     main = "Plant FD vs Frugivore FD",
     xlab = "Plant Functional Dispersion by cell", ylab = "Frugivore Functional Dispersion by cell")
abline(lm(spatial_fdis_frugivore$fdis_frugivore ~ spatial_fdis_plant$fdis_plant))
summary(lm(spatial_fdis_frugivore$fdis_frugivore ~ spatial_fdis_plant$fdis_plant))
```
